---
title: "viz_and_eda"
author: "Shuting Kang"
date: "2022-09-29"
output: html_document
---

## let's import data
```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
```
This code from the course websit

```{r}
## download the dataset online from rnoaa package and only requist Three specific dataset
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
weather_df
```

## Basic scatterplot
```{r}
## for the ggplot (define the dataset we gonna use, and x-axis and y-axis)+add specific geomtry style we require.
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point()
```
the code below could be used instead
```{r}
weather_df %>% 
  ggplot(aes(x=tmin, y=tmax))+
  geom_point()
```

```{r}
weather_df %>% 
  drop_na()%>%
  filter(name=="CentralPark_NY")%>%
  ggplot(aes(x=tmin, y=tmax))+
  geom_point()
```
Let's keep making the same plot but different code and method.
```{r}
weather_scatterplot=
  weather_df %>% 
  drop_na()%>%
  ggplot(aes(x=tmin, y=tmax))

weather_scatterplot+
  geom_point()
```
## advanced scatterplot
```{r}
weather_df%>%
  drop_na()%>%
  ggplot(aes(x=tmin, y=tmax))+
  geom_point(aes(color=name))
```
nice! we get color and have a handly legend. maybe add a smooth curve and make the data points a bit transparent 
```{r}
  ggplot(weather_df,aes(x=tmin,y=tmax))+
  geom_point(aes(color=name),alpha=0.5)+
  geom_smooth(se=FALSE)
```
drop the NA value
```{r}
weather_df%>%
  drop_na()%>%
  ggplot(aes(x=tmin,y=tmax))+
  geom_point(aes(color=name),alpha=0.5)+
  geom_smooth(se=FALSE)
```

the curve gives a sense of the relationship between variables,and the transparency shows where data are overlapping.
Make the saparet plots based on the different dataset
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)
```
drop the na value 
```{r}
weather_df%>%
  drop_na()%>%
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)
```

I've learned a lot about these data. however, yhe relationship between minimum and maximum temperature is little bit boring 
show the relationship between maximum temperature and date(year)
```{r}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp), alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)
```
drop the na value 
```{r}
weather_df%>%
  drop_na()%>%
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp), alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)
```
```{r}
weather_df%>%
  drop_na()%>%
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp), alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)
```

```{r}
weather_df%>%
  drop_na()%>%
  ggplot(aes(x = date, y = tmax, color = prcp)) + 
  geom_point(aes(size = prcp), alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)
```

```{r}
## we adjust the x=axis to make the label easier to visualize 
weather_df%>%
  drop_na()%>%
  ggplot(aes(x = date, y = tmax, color = prcp)) + 
  geom_point(aes(size = prcp), alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)+
  theme(axis.text.x=element_text(angle = 90,vjust=0.5,hjust=1))
```
#### we can compare three variables at the same graph(date, tmax,and prcp)

## Learning assessment
 Write a code chain that starts with weather_df; focuses only on Central Park, converts temperatures to Fahrenheit, makes a scatterplot of min vs. max temperature, and overlays a linear regression line (using options in geom_smooth()).
```{r}
weather_df %>% 
  filter(name == "CentralPark_NY") %>% 
  drop_na()%>%
  mutate(
    tmax_fahr = tmax * (9 / 5) + 32,
    tmin_fahr = tmin * (9 / 5) + 32) %>% 
  ggplot(aes(x = tmin_fahr, y = tmax_fahr)) +
  geom_point(alpha = .5) + 
  geom_smooth(method = "lm", se = FALSE)
 
```
## Odds and ends
```{r}
ggplot(weather_df, aes(x = tmax, y = tmin)) + 
  geom_hex()
```

## Univariable plots
Histograms, barplots, boxplots, violines,...

```{r}
ggplot(weather_df,aes(x=tmax))+
  geom_histogram()
```
```{r}
weather_df%>%
  drop_na()%>%
  ggplot(aes(x=tmax))+
  geom_histogram()
```
```{r}
ggplot(weather_df, aes(x = tmax, fill = name)) + 
  geom_histogram(position = "dodge", binwidth = 2)
```

```{r}
weather_df%>%
  drop_na()%>%
  ggplot(aes(x = tmax, fill = name)) + 
  geom_histogram(position = "dodge", binwidth = 2)
```

```{r}
ggplot(weather_df, aes(x = tmax, fill = name)) + 
  geom_density(alpha = .4, adjust = .5, color = "blue")
```

```{r}
weather_df%>%
  drop_na()%>%
  ggplot(aes(x = tmax, fill = name)) + 
  geom_density(alpha = .4, adjust = .5, color = "blue")
```

### boxplot(Visualize the univariate distribution)
```{r}
ggplot(weather_df, aes(x = name, y = tmax)) + geom_boxplot()
```

```{r}
weather_df%>%
  drop_na()%>%
  ggplot(aes(x = name, y = tmax)) + geom_boxplot()
```

### Violins graph 
```{r}
ggplot(weather_df, aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), alpha = .5) + 
  stat_summary(fun = "median", color = "blue")
```

```{r}
weather_df%>%
  drop_na()%>%
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), alpha = .5) + 
  stat_summary(fun = "median", color = "blue")
```
## Ridge plots were the trendiest plot of 2017, and were replacement for both boxplots and violin plots. They'er implemented in the ggrides package, and are nice if you have a lots of categories in which the shape of the distribution matters.

```{r}
ggplot(weather_df, aes(x = tmax, y = name)) + 
  geom_density_ridges(scale = .85)
```

## Saving and embedding plots

```{r}
weather_plot = ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) 

ggsave("result/weather_plot.pdf", weather_plot, width = 8, height = 5)
```
```{r}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
```

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name))

```